---
title: "linear model from Web"
author: "Diego Martinez"
date: "7/14/22"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS


### scatterplot

```{r}
plot(SIMS~ARM,data=data)
lmr = lm(SIMS~ARM, data = data )
abline(lmr)
```


My scatter plot for SIM~ARMS appears to be correct since there is a positive linear relationship between the two numerical variables. The y-intercept is between -4 and -2. SIMS goes all up to 4. ARM goes all up to 120. The relationship appears to be quite strong based off of this scatter plot. 

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The numerical results conclude with a value of 0.686 (Rounded to the thousandth place). The numerical results were given to me from the table above. 

### Inferential  (Build model.1)

```{r}
summary(lmr)
newdata = c(GRIP = 94, ARM = 88 )
```

The minimum value is -3.6467. The Q1 value is -0.7502. The median is -0.0285. The max is 3.0770. The residual standard error is 1.226 on 145 degrees of freedom. The Multiple R-squared value is 0.4706. The Adjusted R-squared value is 0.467. The p-value is 0.00000000000000022. The p-value is by far lower than .05. 

#### Predict at target point    

```{r}
predictSIM = predict.lm(lmr, newdata = data.frame (GRIP = 94, ARM = 88))
print(predictSIM)
```
Based off of my code, the target point for SIM~ARMS is .706 (Rounded to the thousandth place). I was given this value by the code above. 


#### scatterplot with model fit   
  
```{r}
plot(lmr)
```


I can tell that my scatter plots are accurate because they all relatively have the shape of a straight line. Of course, not perfectly, but this is as accurate as I can get with the model fit. The model fits include Residuals vs. Fitted, Normal Q-Q, Scale Location, and Residuals vs. Leverage.  Plus, at the bottom it shows lm(SIMS~ARM) so I am 100% sure that I'm showing the correct numerical variables for this model. 
  

## Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
Hi = lm(SIMS~GRIP, data = data )
abline(Hi)
```


My scatter plot for SIM~GRIP appears to be correct since there is a positive linear relationship between the two numerical variables. The y-intercept is between -4 and -2. SIMS goes all up to 4. GRIP goes all up to 200. The relationship appears to be quite strong based off of this scatter plot. 


### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

Based off of my code, the numerical value is 0.640 (Rounded to the thousandth place). This number is given to me by the table above. 

### Inferential  (Build model.2)
  
```{r}
summary(Hi)
newdata = c(GRIP = 94, ARM = 88 )
```
The minimum value is -2.9295. The Q1 value is -0.8708. The Q3 value is .8039. The median is -0.1219. The max is 3.3494.  The residual standard error is 1.295 on 145 degrees of freedom. The Multiple R-squared value is 0.4094. The Adjusted R-squared value is 0.4053. The p-value is 0.00000000000000022. The p-value is by far lower than .05. 
 

#### predict model.2 at target point


```{r}
predictGRIP = predict.lm(Hi, newdata = data.frame (GRIP = 94, ARM = 88))
print(predictGRIP)
```
Based off of my coding, the target point value is -0.536 (Rounded to the nearest thousandth).  

#### now add the model fit to our plot for model.2
  
```{r}
plot(Hi)
```

I can tell that my scatter plots are accurate because they all relatively have the shape of a straight line. Of course, not perfectly, but this is as accurate as I can get with the model fit. The model fits include Residuals vs. Fitted, Normal Q-Q, Scale Location, and Residuals vs. Leverage. Plus, at the bottom it shows lm(SIMS~GRIP) so I am 100% sure that I'm showing the correct numerical variables for this model. 


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
data2 = lm(SIMS~ARM+GRIP, data = data)
summary(data2)
```
`
The minimum value is -3.1846. The Q1 value is -0.7380. The Q3 value is .7394. The median is -0.0641. The max is 2.8634. The residual standard error is 1.144 on 144 degrees of freedom.  The Multiple R-squared value is 0.5422. The Adjusted R-squared value is 0.5358. The p-value is 0.00000000000000022. The p-value is by far lower than .05. 

  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point

```{r}
predict.lm(data2, newdata= data.frame(GRIP = 94, ARM = 88))
```

Based off of my code, the target point is .1496 (Rounded to the ten thousandths place). 

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(lmr, data2)
```
The residual degrees of freedom for Model 1 (SIMS~ARM) is 145 while the residual degrees of freedom for Model 2 (SIMS~ARM+GRIP) is 144. There is a difference of 1. The residual sum of squares for Model 1 is 217.88 while the residual sum of squares for Model 2 is 188.43. There is a difference of 29.45 between the residual sum of squares. Model 1 does not have Df, Sum of Sq, F, or Pr(>F). Model 2 has a degrees of freedom of 1. Model 2 has a sum of squares of 29.45. Model 2 has a p-value of 0.000004994. These 2 models differ greatly. 

### Model.2 vs Model.3

```{r}
anova(Hi, data2)
```
The residual degrees of freedom for Model 1 (SIMS~GRIP) is 145 while the residual degrees of freedom for Model 2 (SIMS~ARM+GRIP) is 144. There is a difference of 1. The residual sum of squares for Model 1 is 243.07 while the residual sum of squares for Model 2 is 188.43. There is a difference of 54.64 between the residual sum of squares. Model 1 does not have Df, Sum of Sq, F, or Pr(>F). Model 2 has a degrees of freedom of 1. Model 2 has a sum of squares of 54.64. Model 2 has a p-value of 0.000000001495.  These 2 models differ greatly. 


## Informally compare Model.1 with model.2

```{r}
anova(lmr, Hi)
```
The residual degrees of freedom for Model 1 (SIMS~ARM) is 145 while the residual degrees of freedom for Model 2 (SIMS~GRIP) is 145. There is a difference of 0. The residual sum of squares for Model 1 is 217.88 while the residual sum of squares for Model 2 is 243.07. There is a difference of -25.19 between the residual sum of squares. Model 1 does not have Df, Sum of Sq, F, or Pr(>F). Model 2 has a degrees of freedom of 0. Model 2 has a sum of squares of -25.19. Neither Model 1 or Model 2 have a p-value. These two models are somewhat different. They only share the same residual degrees of freedom and in how they both don't have a p-value. 
